The Role of Context in Spatial Region Identification

### 2013-11-15 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex

%Combine these figures into a single figure across
\begin{figure}
\centering
%% Change the classroom to be a picture stimuli
\begin{tabular}{c c c} 
  \begin{subfigure}[b]{0.32\textwidth}
  \includegraphics[width=\textwidth]{figures/classroom.png}
%  \caption{``The front of the classroom'' could be defined as the area between the desks and the whiteboard where people give presentations.}
  \caption{``The front of the classroom''.}
  \end{subfigure} &
\begin{subfigure}[b]{0.32\textwidth}
  \includegraphics[width=\textwidth]{figures/safety-3.jpg}
%  \caption{Safety for the scout is contextually defined by the location of the friendly turrets and the capabilities of the opposing force.}
  \caption{``The safe region for the scout''.}
\end{subfigure} & 
\begin{subfigure}[b]{0.32\textwidth}
  \includegraphics[width=\textwidth]{figures/bottleneck.JPG}
%  \caption{A bottleneck occurs when paths are confined to a narrow region by geographic features.  In this case, the water prevents ground movement.}
 \caption{``The bottleneck''.}
  \label{fig:safety}
  \end{subfigure}\\
\end{tabular}
  \caption{Examples of spatial regions used as stimuli in our experiments.}
  \label{fig:examples}
\end{figure}

Understanding CDSRs is crucial to joint intelligent action with humans; among other reasons they are frequently tieds to goals and commands: ``go to the front of the classroom'', ``defend the bottleneck'', and ``get to safety''.  We claim that (1) people can effectively communicate using these regions and (2) the size, location and shape of these regions change with the context.  To explore these claims, we present the results of a human study in which people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.  In discussing the results, we consider the applicability of approaches from cognitive systems and robotics.  We posit that to account for this ability, a cognitive system must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.  Current approaches have a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.

@@

The classroom stimuli consisted of pairs of floor plans and photos of classrooms with different chair configurations. For example, Figure \ref{fig:classroom-a} is a floor plan of a classroom where the chairs have been arranged in a circle and Figure \ref{fig:classroom-b} is the accompanying photo of this classroom. We developed four different classroom stimuli each with a different layout of desks: a circular layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli.}. To control for left-right bias effects, two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the images. Consequently, there were seven classroom stimuli (circle, U $\times$ 2, close $\times$ 2, far $\times$ 2).  The two geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the extent of the water.  Finally, the four military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:turrets-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (tanks close to the bottom of the image versus near the top), and the layout of the objects (turrets in a row versus turrets in a row with a lone turret in front).

@@

\textbf{Participants and Procedure:} The experiment was delivered online. 43 undergraduate students from two third-level institutes took part in the study. To control for sequence effects the order of the trials were randomised for each participant as was the order of the stimuli within each trial.

@@

\subsubsection{Predicted Region Changes and Simulation}
\label{sec:simulation}
To assess the claim that changing contexts predictably affects the CDSRs, we identified six expected relationships of region features between stimuli.  For example, we expect the wider bottleneck (g) to result in a larger area than (h), and, in the classroom, the maximum-x coordinate of the region should be greater in the case where the desks are further away from the whiteboard (f) than (e).  CDSRs are dynamic, in the move to safety scenario, we expect that people will simulate the actions of the agents.  Therefore, we expect the maximum-Y to be larger in the tanks scenario (j) than the turrets (k) because the friendly tanks will pursue the enemy tanks.  Similarly, we expect maximum-Y coordinate to the be same in the two different turrets scenarios (k) and (l), because people would simulate the results of the battle and realize that a single turret will not provide safety against three tanks.

@@

Table \ref{tab:region-location-tests} compares the expected relationships between the CDSR features between stimuli.  In five of the six comparisons, the expected relationships were found, supporting our claim that changes of context systematically change the perceived extent of a CDSR.  The last two rows test our hypothesis that people are simulating the actions of others in defining the extent of the CDSR.  The comparison between the maximum-y coordinate of the safety area with tanks (j) versus turrets (k) provides some evidence, although not statistically significant, that people are simulating actions of other agents in the scene.  In the other simulation case, we did not observe the expected relationship between the row of turrets (k) and turrets +1(l).  The average maximum-Y was greater for the turret +1 stimuli indicating that people felt that the single turret provided protection from the enemy tanks.  As discussed in the previous section, we expected some ambiguity in this trial as people likely have different models for the capabilities of the different units in this scenario.

@@

To assess the claim that changing contexts predictably affects the CDSRs, we can compare the ratings of points between stimuli.  In the classroom domain, we assume that part of the definition of the front of the room is that where one would stand when teaching a class, therefore, we expected the rating for the center point, P2, to be higher in the u-shaped scenario (c) than the rows scenarios (e) and (f).  This is confirmed in the results from Table \ref{tab:point-likert-scores}.  In the military domain,  In stimulus (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimulus (j) the P2 point is in \textit{no-man's land}\footnote{We use the term no-man's land to describe the unoccupied and contested area between the opposing forces.}.  Given this change in context one would expect a higher ranking for the P2 point in context (i) than in context (j). This is confirmed in Table \ref{tab:point-likert-scores} and is statistically significant $(p<.05)$.  To assess the extent of participants use of simulation in determining the context, we compare the scores the \textit{no-man's land} point between the tanks (j) to the turrets (k) stimuli and the turrent (k) and turrets +1 (l).  The mobility of the friendly tanks should increase the size of the safe area over time resulting in higher ratings in stimuli (j) than (k).  The single turret should offer only temporary protection against the energy tank force, therefore it should not increase the safety rating.  Neither of these results were found in Table \ref{tab:point-likert-scores}.  As discussed in Section \ref{sec:simulation}, we expected ambiguity with the turrets +1 stimuli.  The difference in the rating of P2 between the tanks and turrets scenario provides two alternative explanations regarding how participants interpreted the context:  (1) the turrets were viewed as superior defensive units and, thus, better able to provide safety, or (2) the turrets being fixed could not move away, and thus, establish a more permanent CDSR for safety.  Further experiments are needed to explore these hypotheses.

@@

\subsection{discussion}
Taken together, these results support the following conclusions. First, people are able to agree on location and area of CDSRs.  This agreement is reduced in situations where the domain knowledge of the context is ambiguous, such as, the circular desk arrangement or the defensive capabilities of a single turret.  Second, changing the context predictably changes the subjects assessments of location, shape and size of the CDSR.  These results were supported by tasks involving people labeling regions as well as rating individual points.  On the other hand, our experiments produced mixed results concerning our hypothesis that people use simulation to evaluate dynamic changes to CDSRs.  Further studies are required to control for the domain knowledge of the participants exploring this question.

@@

Whilst spatial reasoning has been an active topic in AI and robotics for decades, the properties we have demonstrated for CDSRs (i.e. that they vary systematically with the function and configuration of objects, plus the agents interacting with them) are not addressed by previous work, and thus it is not possible to take existing work and build a cognitive system capable of reasoning about, or communicating with, CDSRs. The remainder of this section reviews related previous work with reference to the requirements of modelling CDSRs.

>>>>

%Combine these figures into a single figure across
\begin{figure}
\centering
%% Change the classroom to be a picture stimuli
\begin{tabular}{c c c} 
  \begin{subfigure}[b]{0.32\textwidth}
  \includegraphics[width=\textwidth]{figures/classroom-label.png}
%  \caption{``The front of the classroom'' could be defined as the area between the desks and the whiteboard where people give presentations.}
  \caption{``The front of the classroom''.}
  \end{subfigure} &
\begin{subfigure}[b]{0.32\textwidth}
  \includegraphics[width=\textwidth]{figures/tanks-close-label.jpg}
%  \caption{Safety for the scout is contextually defined by the location of the friendly turrets and the capabilities of the opposing force.}
  \caption{``The safe region for the scout''.}
\end{subfigure} & 
\begin{subfigure}[b]{0.32\textwidth}
  \includegraphics[width=\textwidth]{figures/bottleneck-label.JPG}
%  \caption{A bottleneck occurs when paths are confined to a narrow region by geographic features.  In this case, the water prevents ground movement.}
 \caption{``The bottleneck''.}
  \label{fig:safety}
  \end{subfigure}\\
\end{tabular}
  \caption{Examples of the spatial regions used as stimuli in our experiments.}
  \label{fig:examples}
\end{figure}

Understanding CDSRs is crucial to joint intelligent action with humans; among other reasons they are frequently tied to goals and commands: ``go to the front of the classroom'', ``defend the bottleneck'', and ``get to safety''.  We claim that (1) people can effectively communicate using these regions and (2) the size, location and shape of these regions change with the context.  To explore these claims, we present the results of a human study in which people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.  In discussing the results, we consider the applicability of approaches from cognitive systems and robotics.  We posit that to account for this ability, a cognitive system must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.  Current approaches have a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.

@@

The classroom stimuli consisted of pairs of floor plans and photos of classrooms with different chair configurations. For example, Figure \ref{fig:classroom-a} is a floor plan of a classroom where the chairs have been arranged in a circle and Figure \ref{fig:classroom-b} is the accompanying photo of this classroom. We developed four different classroom stimuli each with a different layout of desks: a circular layout, a U-shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli.}. To control for left-right bias effects, two versions of the U-shaped layout (c), close rows layout (e) and far rows layout (f) were created by horizontally flipping the images. Consequently, there were seven classroom stimuli (circle, U $\times$ 2, close $\times$ 2, far $\times$ 2).  The two geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the extent of the water.  Finally, the four military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:turrets-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (tanks close to the bottom of the image versus near the top), and the layout of the objects (turrets in a row versus turrets in a row with a lone turret in front).

@@

\textbf{Participants and Procedure:} The experiment was delivered online. Forty-three undergraduate students from two third-level institutes took part in the study. To control for sequence effects the order of the trials were randomised for each participant as was the order of the stimuli within each trial.

@@

\subsubsection{Predicted Region Changes and Simulation}
\label{sec:simulation}
To assess the claim that changing contexts predictably affects the CDSRs, we identified six expected relationships of region features between stimuli.  For example, we expect the wider bottleneck (g) to result in a larger area than (h), and, in the classroom, the maximum-x coordinate of the region should be greater in the case where the desks are further away from the whiteboard (f) than (e).  CDSRs can be dynamic. Thus, in the move to safety scenario, we expect that people will simulate the actions of the agents.  Therefore, we expect the maximum-Y to be larger in the tanks scenario (j) than the turrets (k) because the friendly tanks will pursue the enemy tanks.  Similarly, we expect maximum-Y coordinate to the be same in the two different turrets scenarios (k) and (l), because people would simulate the results of the battle and realize that a single turret will not provide safety against three tanks.

@@

Table \ref{tab:region-location-tests} compares the expected relationships between the CDSR features between stimuli.  In five of the six comparisons, the expected relationships were found, supporting our claim that changes of context systematically affect the perceived extent of a CDSR.  The last two rows of the table evaluate our hypothesis that people are simulating the actions of others in defining the extent of the CDSR.  The comparison between the maximum-y coordinate of the safety area with tanks (j) versus turrets (k) provides some evidence, although not statistically significant, that subjects were simulating actions of other agents in the scene.  In the last row, we did not observe the expected relationship between the row of turrets (k) and turrets +1(l).  The average maximum-Y was greater for the turret +1 stimuli indicating that people felt that the single turret provided protection from the enemy tanks.  As discussed in the previous section, we expected some ambiguity in this trial as people likely have different models for the capabilities of the different units in this scenario.

@@

To assess the claim that changing contexts predictably affects the CDSRs, we can compare the ratings of points between stimuli.  In the classroom domain, we assume that part of the definition of the front of the room is where one would stand when teaching a class, therefore, we expected the rating for the center point, P2, to be higher in the U-shaped scenario (c) than the rows scenarios (e) and (f).  This is confirmed in the results from Table \ref{tab:point-likert-scores}.  In the military domain, in stimulus (i), the friendly tanks are positioned between the P2 point and the enemy tanks, whereas, in stimulus (j), the P2 point is in \textit{no-man's land}\footnote{We use the term no-man's land to describe the unoccupied and contested area between the opposing forces.}.  Given this change in context one would expect a higher ranking for the P2 point in context (i) than in context (j). This is confirmed in Table \ref{tab:point-likert-scores} and is statistically significant $(p<.05)$.  To assess the extent of participants use of simulation in determining the context, we compare the subject's ratings of the \textit{no-man's land} point between the tanks (j) and the turrets (k) stimuli as well as between the turrent (k) and turrets +1 (l).  The mobility of the friendly tanks should increase the size of the safe area over time resulting in higher ratings in stimuli (j) than (k).  The single turret should offer only temporary protection against the enemy tank force, therefore it should not increase the safety rating.  Neither of these results were found in Table \ref{tab:point-likert-scores}.  As discussed in Section \ref{sec:simulation}, we expected ambiguity with the turrets +1 stimuli.  The difference in the rating of P2 between the tanks and turrets scenario provides two alternative explanations regarding how participants interpreted the context:  (1) the turrets were viewed as superior defensive units and, thus, better able to provide safety, or (2) the turrets being fixed could not move away, and thus, establish a more permanent CDSR for safety.  Further experiments are needed to explore these hypotheses.

@@

\subsection{Discussion}
Taken together, these results support the following conclusions. First, people are able to agree on location and area of CDSRs.  This agreement is reduced in situations where the domain knowledge of the context is ambiguous, such as, the circular desk arrangement or the defensive capabilities of a single turret.  Second, changing the context predictably changes the subjects assessments of location, shape, and size of the CDSR.  These results were supported by tasks involving people labeling regions as well as rating individual points.  On the other hand, our experiments produced mixed results concerning our hypothesis that people use simulation to evaluate dynamic changes to CDSRs.  Further studies are required to control for the domain knowledge of the participants exploring this question.

@@

Whilst spatial reasoning has been an active topic in AI and robotics for decades, the properties we have demonstrated for CDSRs (i.e. that they vary systematically with the function and configuration of objects, plus the agents interacting with them) are not addressed by previous work.  Thus it is not possible to take existing work and build a cognitive system capable of reasoning about, or communicating with, CDSRs.  The remainder of this section reviews related previous work with reference to the requirements of modelling CDSRs.
